EEG characterization and classification based on the histograms of visual gradients of plots

### 2016-12-18 article.tex
The paper is materializing.....

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
$^{1}$ \quad Computer Engineering Department, Instituto Tecnológico de Buenos Aires ITBA; info@itba.edu.ar}

@@

% Current address and/or shared authorship
\firstnote{Current address: Lavarden 315, Postal Code C1437FBH, City of Buenos Aires, Argentina} 
\secondnote{These authors contributed equally to this work.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}
The template details the sections that can be used in a manuscript. Note that the order of article sections may differ from the requirements of the journal (e.g. for the positioning of the Materials and Methods section). Please check the instructions for authors page of the journal to verify the correct order. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX related questions please contact Janine Daum at latex-support@mdpi.com.

One of the most shoking reasons is that we can use the capacity of ML techniques to automatically explore the signal, by using these new featrues.

@@

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. And now citing a book reference \cite{ref-book}.

In this work EEG signals from ALS patients are offline processed with this radically different and horizontal approach.  This is the continuation of the work presented here and here.

The same methodological algorithm can be used to process rhythmic EEG markers or paradigms and at the same time to analyze spatiotemporal phenomena like ERP, particularly P300.

@@

@@

@@

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.

@@

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
entry 1		& data			& data\\
entry 2		& data			& data\\
\bottomrule
\end{tabular}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{All sources of funding of the study should be disclosed. Please clearly indicate grants that you have received in support of your research work. Clearly state if you received funds for covering the costs to publish in open access.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``X.X. and Y.Y. conceived and designed the experiments; X.X. performed the experiments; X.X. and Y.Y. analyzed the data; W.W. contributed reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be limited to those who have contributed substantially to the work reported.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictofinterests{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funding sponsors in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript, or in the decision to publish the results must be declared in this section. If there is no role, please state ``The founding sponsors had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript, and in the decision to publish the results''.}

@@

%=====================================
% References, variant A: internal bibliography
%=====================================
\renewcommand\bibname{References}
\begin{thebibliography}{999}
% Reference 1
\bibitem{ref-journal}
Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
% Reference 2
\bibitem{ref-book}
Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
\end{thebibliography}

%=====================================
% References, variant B: external bibliography
%=====================================
%\bibliography{your_external_BibTeX_file}

>>>>

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{%
$^{1}$ \quad Computer Engineering Department, Instituto Tecnológico de Buenos Aires (ITBA); info@itba.edu.ar}

@@

% Current address and/or shared authorship
\firstnote{Current address: C1437FBH Lavarden 315, Ciudad Autónoma de Buenos Aires, Argentina} 
\secondnote{These authors contributed equally to this work.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.

@@

\begin{enumerate}[leftmargin=*,labelsep=3mm]
\item	Why
\item	Purpose and significance
\item	Current state of the research field and key publications cited
\item   Controversial hypothesis
\item   Main aim and principal conclusions.
\end{enumerate}

Cybathlon

x) El paper tiene un foco general en EEG y rapidamente se centra en BCI como una prueba del método.
x) La relevancia de estudiar métodos de análisis de EEG para pacientes con ALS.
x) Ofrece una Receta de cómo implementar un esquema simple de detección de P300 y ofrece una comparativa contra el paper de las italianas.
x) La transversalidad del método de poder detectar eventos rítmicos asi como transientes como p300.
x) La posibilidad de utilizarlo para reconocer, fuertemente un patron particular y con la flexibilidad de detectar variantes basado en las poblaciones de descriptores (esta es mi hipótesis de porque funciona, porque yo vi que el método se hace pelota con pequeñas variantes de la señal).
x) La importancia del averaging de EEG y sus desbalanceos, puntualizando que en el paper original
x) Que la caída en la clasificación no cae significativamente aún bajando la cantidad de repeticiones requeridas para el averaging.
x) ¿ Por qué da mejor en PO8 que en Cz y Pz que son los canales P300 ?  (como future work).

Although recently impressive advances in neuroimagining techniques has pushed EEG, Electroencephalography technology a little bit out of the way in clinical context, where it has always being used since its discovery by Berger.

Their versatility, ease of use, temporal resolution, ease and development of fabrication, and the explosion as a consumer devices, is pushing EEG devices to become the de facto non invasive method to access and harness brain information.

Key contribution to this development has been the field of Brain Computer Interfaces, which is the development of a new channel of communication particularly aimed to person with some degree of neuro degenerative disease.

One very remarkable aspect of this communication channel is the ability to transmit some general cognitive state, like alertness, drowsiness, boredom, and so on, which can very helpful particularly in rehabilitation procedures. \citep{Guger2009a}.

One of the most shocking reasons is that we can use the capacity of ML techniques to automatically explore the signal, by using these new features.

Additionally, they have a high variability between different subjects and even between different moments for the same subject.  This inherent complexity comes from  is a real challenge when it is required to extract information from raw EEG signals.

Signal is complex and it is difficult to find those values which are not variable.

Due to this inner complexity, it is often necessary to implement many distinct and specialized algorithmic methods, to filter the signal, classify it, and try to determine some meaning out of it.

In this work EEG signals from ALS patients are offline processed with this radically different and horizontal approach.  This is the continuation of the work presented here and here \citep{Ramele2016}, which was previously applied to rythmic patterns.

The same methodological algorithm can be used to process rhythmic EEG markers or paradigms and at the same time to analyze spatiotemporal phenomena like ERP, particularly P300 \citep{Ramele2016}.

@@

Successful applications of BCI has been applied to rehabilitation \citep{Jure2016}.

This paper reports on a study done to address
the hypothesis that communication task complexity
(copy-spelling vs. novel text generation of a picture
description) is correlated with BCI performance accuracy
and P300 characteristics.

BCI accuracy
was calculated as the percentage of attempted selections
that matched the intended selection.

\section{Materials and Methods}

Besides new novel applications of BCI, the core value of the entire discipline is to provide communication assistance to patients.  It is understood that people suffering from ALS, even though the disease do not harm the CNS, their brain signals and suffer slow and deterioriating changes that lead to more complex patterns to be detected [cite]

Although they are used (PAPER DE P300 donde dice que se usa en los laboratorios) they still lack the bla bla bla CITE LOTTE BOOK \citep{Huggins2016}

We use the dataset 008-2014 published on the BNCI-Horizon website.  This consortium of companies and universities were developed under the umbrella of the EU to define, provide guidance for BCI development \citep{Riccio2013}.

\subsection{Experimental protocol}

The protocol is explained [HERE] but can be summarized as follows:  8 subjects affected by different levels of ALS disease, were used to generate all the data. The P300 task was meant to spell 7 words of 5 letter each.  The first 3 runs were used for training and the remaining 4 for testing with visual feedback.  A trial, for this experiment, was defined as each attempt to select a letter from the speller, and it was composed of 10 repetions of flashes of 6 rows and 6 columns of the matrix, yielding 120 repetitions.  Flashing of row/columns was performed for 0.125 s, following by a resting period of the same size.

This dataset was recorded from a healthy subject during a no-feedback session. She/he sat in a normal chair with relaxed arms resting on the table and fingers in the standard typing position at the computer keyboard. The task was to press with the index and little fingers the corresponding keys in a self-chosen order and timing 'self-paced key typing'. The experiment consisted of 3 sessions of 6 minutes each. All sessions were conducted on the same day with some minutes break in-between. Typing was done at an average speed of 1 key per second.

\subsection{Morphological Signal Analysis}

Due to the inner complexity of EEG signals, several approaches and methods has been applied to decode them.  Of particular interests are morphological signal analysis 
Paper del chabon del ITBA que detalla bien los diferentes metodos,

paper de los mexicanos \citep{Alvarado-Gonzalez2016}.

Morphological or

template matching

spike
sharp wave
spike wave
polyspike 
slow wave complex

shape domain

\subsubsection{P300}

Uhhhh I can talk a lot about P300 \citep{Knuth2006}.

@@

\subsection{Algorithm}

\subsubsection{Preprocessing}

\subsubsection{Signal to Image Transformation}

There are many ways an EEG matrix can be transformed into an image.

What we do is take the signal and plot it.  Thats all folks.

\subsubsection{Signal Averaging}

The methods for averaging.

\subsubsection{Classification}

Finally the classification is carried out by using the NBNN algorithm which can be described as follows.

BCI Simulation vs k-fold cross validation

@@

Results are shown in the following chart.

\begin{figure}[H]
\centering
\includegraphics[width=18cm]{subjectaverageds4.eps}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

Contrary to the literature (\citep{Huggins2016,Jure2016}) the best performing channels were not Cz and Pz and instead Pz.

GRAND AVERAGES

it is quite likely that the data will be unbalanced (that is, an unequal number of trials in each condition)

\begin{table}[H]
\caption{Results obtained for Accuracy levels by k-fold cross validation.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{cccc}
\toprule
\textbf{Cz}	& \textbf{Cz}	& \textbf{Best Channel}	& \textbf{Performance}\\
\midrule
0.845000     & 0.794286 & 2 & 0.794286 $\pm$ 0.011952 \\
0.863000     & 0.811429 & 7 & 0.934286 $\pm$ 0.011200 \\
0.872000     & 0.780000 & 2 & 0.780000 $\pm$ 0.031843 \\
0.859000     & 0.675714 & 8 & 0.904286 $\pm$ 0.013232 \\
0.862000     & 0.801429 & 7 & 0.932857 $\pm$ 0.008400 \\
0.886000     & 0.964286 & 2 & 0.964286 $\pm$ 0.010435 \\
0.886000     & 0.784286 & 7 & 0.930000 $\pm$ 0.009744 \\
0.923000     & 0.908571 & 7 & 0.997143 $\pm$ 0.003733 \\
\bottomrule
\end{tabular}
\end{table}

The ITR, or BTR, in the case of reactive BCIs (CITE WOLPAW) strongly depends on the amount of averaging required to transmit a valid and robust selection.  If the amount of repetitions is reduced for each subject, performing a BCI Simulation can be shown on Fig. \ref{fig:singletrial}

\begin{figure}[H]
\centering
\includegraphics[width=18cm]{singletriality.eps}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\label{fig:singletrial}
\end{figure}

This method, different from other methods which is based on the nonlinearity of the gradient of histograms which can be used to detect

is also based on how the image look like.

Results of one against the other

SNR of p300 and how to detect it

Check if you can use this to detect any kind of transient signbal

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

With this result we showed a method to process EEG signals where the main characteristic can be both, rhytmic in nature as in motor imagery, and also transient and in time space, like the P300 event related potential.

If we avoid removing artefacts achieve a higher value of classification.

Si el desempeño da parecido hay que escribirlo ya que con menos valores se permite un ITR mayor.

We also achieved single trial classification for just one subject.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{Juan Miguel Santos organized the project and conceived the idea. All the authors equally worked in the completion of this project.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictofinterests{The authors declare no conflict of interest.}

@@

%=====================================
% References, variant A: internal bibliography
%=====================================
%\renewcommand\bibname{References}
%\begin{thebibliography}{999}
% Reference 1
%\bibitem{ref-journal}
%Lastname, F.; Author, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142-149.
% Reference 2
%\bibitem{ref-book}
%Lastname, F.F.; Author, T. The title of the cited contribution. In {\em The Book Title}; Editor, F., Meditor, A., Eds.; Publishing House: City, Country, 2007; pp. 32-58.
%\end{thebibliography}

%=====================================
% References, variant B: external bibliography
%=====================================
\bibliography{article}
